% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 01-matlab-at-a-glance.tex -- getting on your feet in MATLAB
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Lesson 01: MATLAB at a Glance}
\input{postamble}

\begin{document}
\renderTitle

\noindent
MATLAB can be unbelievably \emph{frustrating...}

\

\noindent
I have lost count of the number of times this program has caused me
immeasurable amounts of pain through no fault of my own. With poor
language design, undocumented quirks, and, at times, seemingly
non-deterministic behavior, MATLAB is a beast to tame.  It will
shamelessly stab you when you least expect it and show you absolutely
zero remorse!

\

\noindent
That said, it can be incredibly useful when the stars align, so here we
are...

\section{The MATLAB Environment}

Before we jump head first into the fundamentals of MATLAB, it helps if
we take a step back and go over how to interact with its environment.

\subsection{So What’s an M-File?}

\vocab{M-files}, denoted with a \texttt{.m} extension, are the primary
method for scripting MATLAB operations.  MATLAB executes the contents of
these files as if you had typed them explicitly into the command prompt.
Confusingly, M-files can be both executable user scripts or define
functions, which we can then reference in other M-files.  To execute the
contents of an M-file, type its name into the command prompt without the
\texttt{.m} extension.

\subsection{But Where to Find M-Files?}

\vocab{The MATLAB Search Path}\footurl{https://www.mathworks.com/help/%
matlab/matlab_env/what-is-the-matlab-search-path.html} specifies the
paths MATLAB will search for M-files.  It will first search the current
working directory (accessible with \mCommand[https://www.mathworks.com/%
help/matlab/ref/pwd.html]{pwd}), your \mCommand[https://www.mathworks%
.com/help/matlab/ref/userpath.html]{userpath}, paths specified in the
\texttt{MATLABPATH} environment variable, and finally, paths added by
installed packages.  To view the paths in the current search path, you
can use the \mCommand[https://www.mathworks.com/help/matlab/ref/path.%
html]{path} command.  To locate the path of an M-file, use the
\mCommand[https://www.mathworks.com/help/matlab/ref/which.html]{which}
command followed by the name of the M-file you're trying to find.

\subsection{User-Defined Startup Script}

If you'd like to execute some commands before MATLAB starts, you can add
a \mCommand[https://www.mathworks.com/help/matlab/ref/startup.html]%
{startup} M-file in the search path.  This file is handy if you wish to
modify the environment at runtime (i.e. set graphic options).  Feel free
to take a look at my \href{https://github.com/jacobkoziej/dotfiles/%
blob/master/.config/matlab/startup.m}{\texttt{startup.m}} if you'd like
to see an example.

\section{Fundamentals}

Now that we have the \emph{slightest} idea of what's going on, it's out
the frying pan and into the fire!

\subsection{Everything's a Matrix!}

Who would have thought that MATrix LABratory represents its data as
matrices?  Matrices generally have three forms:

\begin{itemize}
	\item
		Matrices of the form \(\matrixField{m}{n}{F}\)

	\item
		Vectors of the form \(\matrixField{m}{1}{F}\) or
		\(\matrixField{1}{n}{F}\)

	\item
		Scalars of the form \(\matrixField{1}{1}{F}\)
\end{itemize}

\noindent
Certain MATLAB commands may want a specific type of matrix (i.e. a row
vector) or may operate on different portions of a matrix (i.e. its
columns).  We will see examples of this later.  For now, let's go over
how you might instantiate different types of matrices:\aside{Well,
actually, these are all represented as N-dimensional arrays with an
infinite amount of singleton dimensions.\footnotemark}\footnotetext{%
\url{https://blogs.mathworks.com/loren/2006/08/09/essence-of-indexing/}}

\inputminted{matlab}{01-matlab-at-a-glance.d/instantiate-matrix.m}

\subsection{Array \& Matrix Operations}

Array operations execute on a per-element basis, while matrix operations obey
the rules of linear algebra.

\begin{longtable}{p{0.15\textwidth}p{0.4\textwidth}}
\toprule
Operator & Purpose \\
\midrule
\mCommand{A + B}  & Addition                    \\
\mCommand{+A}     & Unary addition              \\
\mCommand{A - B}  & Subtraction                 \\
\mCommand{-A}     & Unary subtraction           \\
\mCommand{A .* B} & Element-wise multiplication \\
\mCommand{A .^ B} & Element-wise power          \\
\mCommand{A ./ B} & Right array division        \\
\mCommand{A .\ B} & Left array division         \\
\mCommand{A.'}    & Array transpose             \\
\bottomrule
\\
\caption{Array Operations}
\end{longtable}

\begin{longtable}{p{0.15\textwidth}p{0.4\textwidth}}
\toprule
Operator & Purpose \\
\midrule
\mCommand{A * B}  & Multiplication \\
\mCommand{A \ B}  & Left division  \\
\mCommand{A / B}  & Right division \\
\mCommand{A ^ B}  & Power          \\
\bottomrule
\\
\caption{Matrix Operations}
\end{longtable}

\subsection{Memory Representation}

MATLAB stores matrices in column-major order.  Each entry can be real or
complex-valued.  In the case of a matrix consisting of complex values,
MATLAB will internally allocate two matrices: one for the real component
of an entry and another for the complex component of an entry.  Although
you don't see this duplicate matrix business on your end, it is
important to keep this in mind as each element of these matrices is an
IEEE 754 double\footurl{https://en.wikipedia.org/wiki/IEEE_754}, which
can use a lot of memory as matrices grow in size!

\subsection{Sparse Matrices}

\vocab{Sparse matrices}\footurl{https://en.wikipedia.org/wiki/Sparse%
_matrix} are matrices that consist mostly of zero entries.  Although
they \emph{can} be represented as regular matrices, once they grow to
insane sizes (i.e.~\(\matrixField{1,000,000}{1,000,000}{F}\)), the walls
of the real world start crashing down.  To get around this, MATLAB
matrices either have a full or sparse \vocab{storage class}, the latter
made for efficiently storing sparse data.

MathWorks has worked hard for full and sparse matrices to play nicely,
and for the most part, they do as almost any operation that works on a
full matrix works for a sparse matrix.  With the exception of a few
operations, any operation performed with a sparse matrix will result in
a sparse matrix.

To learn more about sparse matrices, visit \url{https://www.mathworks.%
com/help/matlab/sparse-matrices.html}, and if you're interested in their
implementation, feel free to read \url{https://www.mathworks.com/help/%
pdf_doc/otherdocs/simax.pdf}.
\end{document}
