% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 02-vectorization.tex -- writing performant MATLAB code
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Lesson 02: Vectorization}
\input{postamble}

\begin{document}
\renderTitle

So far, what we've learned hasn't been all that useful.  Sure, we can
instantiate matrices, but we're here to make MATLAB work like a horse.
One of the fundamental principles behind \emph{clean \& efficient} code
is the vectorization of operations.  By vectorizing, our code appears
more like mathematical expressions found in textbooks, introduces fewer
opportunities for errors, and runs much faster as it can utilize
hardware vector operations!\footurl{https://www.mathworks.com/help/%
matlab/matlab_prog/vectorization.html}

\section{Vector Generation}

Say we want to create a vector \(\mat{x} = [i : i \in \Z_8]\).  We could
do it like so:

\mCommandOutput{02-vectorization.d/basic-vector}

We achieved the following result by utilizing the \mCommand[https://www%
.mathworks.com/help/matlab/ref/colon.html]{:} operator.  This operator
is useful for generating vectors, which we can use either for
calculations or indexing into arrays.

The operator also allows us to specify a step increment other than one:

\mCommandOutput{02-vectorization.d/step-vector}

But what if we want to create a vector over some range with a fixed
amount of points (i.e. four evenly spaced values for \(t \in [0, 1]\))?
We could instead utilize \mCommand[https://www.mathworks.com/help/%
matlab/ref/linspace.html]{linspace()}:

\mCommandOutput{02-vectorization.d/linspace}

\section{Indexing}

Now, to \emph{ruin} your day.  MATLAB starts indexing at one...

Let's reuse \(\mat{x}\) from the previous section.  We can access the
third element of \(\mat{x}\) like so:

\mCommandOutput{02-vectorization.d/third-element}

We can also access multiple elements at once:

\mCommandOutput{02-vectorization.d/multiple-elements}

Or if we're so inclined, the penultimate element:

\mCommandOutput{02-vectorization.d/penultimate-element}

Matrices logically follow by adding another index:

\mCommandOutput{02-vectorization.d/matrix-index}

\begin{minipage}{\textwidth}
	We can also index \emph{linearly:}

	\vspace{1em}

	\mCommandOutput{02-vectorization.d/linear-index}
\end{minipage}

But that's an adventure for another time \texttt{;)}

\section{Size \& Shape}

When dealing with N-dimensional arrays, it's important to know how to
get basic information related to the size and shape.  Lucky for us,
MATLAB offers a few convenient functions just for this!

\begin{itemize}
	\item
		\mCommand[https://www.mathworks.com/help/matlab/ref/%
		length.html]{length()} --- the length of the largest
		array dimension

	\item
		\mCommand[https://www.mathworks.com/help/matlab/ref/%
		size.html]{size()} --- the size of the array

	\item
		\mCommand[https://www.mathworks.com/help/matlab/ref/%
		ndims.html]{ndims()} --- the number of array dimensions

	\item
		\mCommand[https://www.mathworks.com/help/matlab/ref/%
		numel.html]{numel()} --- the number of array elements
\end{itemize}

\section{Reshaping Arrays}

Say we wanted to create the following matrix:

\begin{equation}
	\mat{A}
	=
	\begin{bmatrix}
		1 & 4 & 7\\
		2 & 5 & 8\\
		3 & 6 & 9\\
	\end{bmatrix}
\end{equation}

We can agree that it's nothing more than

\begin{equation}
	\mat{a}
	=
	[1, 2, 3, 4, 5, 6, 7, 8, 9],
\end{equation}

where each column of \(\mat{A}\) is every three elements of \(\mat{a}\).

\begin{minipage}{\textwidth}
	We can achieve this pragmatically with \mCommand[https://www.%
	mathworks.com/help/matlab/ref/reshape.html]{reshape()}:

	\vspace{1em}

	\mCommandOutput{02-vectorization.d/reshape}
\end{minipage}

What's important to note is that for a reshape to be possible, the
resultant array must have the same number of elements.  Given this, we
can let MATLAB do the hard work of determining a dimension for us:

\mCommandOutput{02-vectorization.d/lazy-reshape}

We could also \emph{flatten} a matrix by reshaping, but you'll rarely
see a MATLAB programmer do it this way.  Instead, most programmers will
linearly index all the elements of the matrix:

\mCommandOutput{02-vectorization.d/flatten}

\section{Broadcasting}

Probably the most trippy thing when you first start using MATLAB is the
concept of \vocab{broadcasting}.  In essence, when working with array
operations, MATLAB will take the liberty of \emph{implicitly} doing a
\mCommand[https://www.mathworks.com/help/matlab/ref/repmat.html]%
{repmat()} to make the sizes of operations work out.

\begin{minipage}{\textwidth}
	Take for example:

	\vspace{1em}

	\mCommandOutput{02-vectorization.d/broadcast}
\end{minipage}

Strange, but it's as if we had performed the following operation
manually:

\inputminted{matlab}{02-vectorization.d/broadcast-implicit.m}

Although broadcasting is \emph{very} convenient, it is always waiting
for you to drop your guard to stab you in the back when you least expect
it.  Having been a victim of MATLAB's senseless violence, you can do
nothing besides being cautious around broadcasting operations to avoid
introducing some \emph{nasty} bugs.\footnote{Worst of all, you might
even introduce a broadcasting operation unintentionally; that's when
the real head-banging can begin!}

\section{Function Calls}

Depending on which function you call, its behavior on matrices will be
different.  Let's reuse matrix \(\mat{A}\) from the previous section:

\mCommandOutput{02-vectorization.d/exp}

As you can see, \mCommand[https://www.mathworks.com/help/matlab/ref/%
exp.html]{exp()} returns the exponential evaluated on each element of an
array. However, this is not always the case; take, for example:

\mCommandOutput{02-vectorization.d/sum}

Notice that \mCommand[https://www.mathworks.com/help/matlab/ref/sum.%
html]{sum()} only evaluated the sum of each column of \(\mat{A}\). This
example illustrates why it's important to read a function's
documentation \emph{before} you use it, as it may not behave as you'd
expect!

Now, we turn to an interesting experiment where we sum all the elements
of a matrix. Let's generate a \enquote{large} matrix with uniformly
distributed pseudorandom integers:

\inputminted{matlab}{02-vectorization.d/randi.m}

Now, let's compare the performance of summing all the elements of matrix
\(\mat{B}\) using two methods:

\mCommandOutput{02-vectorization.d/sum-flattened}

\mCommandOutput{02-vectorization.d/sum-all}

Wow, I don't know about you, but flattening is \emph{significantly}
slower than just calling \mCommand[https://www.mathworks.com/help/%
matlab/ref/sum.html]{sum()} correctly.  The reason for this dramatic
difference is that we need to make a flattened \emph{copy} of
\(\mat{B}\) in memory before we can sum all the values.

The moral of the story is that a bit of reading can save you a lot of
time!
\end{document}
