% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 03-indexing.tex -- getting what you want
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Lesson 03: Indexing}
\input{postamble}

\begin{document}
\renderTitle

At first, indexing in MATLAB seems like a trivial topic, and for the
most part, the basics will suffice for most \emph{basic} operations.
However, once you start trying to vectorize more complicated operations,
especially ones that rely on conditions, you'll quickly find yourself
limited in what you can calculate.

\section{Subscript Indexing}

For the remainder of this lesson, we'll be working with the following:
\begin{equation}
	\mat{x}
	=
	[1, 2, 3, 4, 5, 6, 7, 8]
\end{equation}
\begin{equation}
	\mat{A}
	=
	\begin{bmatrix}
		1 & 5 &  9 & 13 \\
		2 & 6 & 10 & 14 \\
		3 & 7 & 11 & 15 \\
		4 & 8 & 12 & 16 \\
	\end{bmatrix}
\end{equation}

We've already seen that we can access multiple elements of a vector by
indexing with another vector:

\mCommandOutput{03-indexing.d/seen}

But we can also write to these elements:

\mCommandOutput{03-indexing.d/write}

Matrices behave much the same as this behavior extends to each
dimension:

\mCommandOutput{03-indexing.d/matrix}

Say we want to select all elements of a dimension; we can use a bare
\mCommand{:} in place of \mCommand{1:end} to save us some typing:

\mCommandOutput{03-indexing.d/col}

\section{Linear Indexing}

Last lesson, we flirted with the idea of linear indexing without getting
too much into how it works.  Essentially, MATLAB stores its arrays in
column-major order,  and linear indexing is nothing more than taking
advantage of this design decision.

Take for example:

\mCommandOutput{03-indexing.d/linear}

You might find yourself having to convert between subscripts and linear
indices depending on what you're trying to achieve, and for that MATLAB
has the \mCommand[https://www.mathworks.com/help/matlab/ref/sub2ind.%
html]{sub2ind()} and \mCommand[https://www.mathworks.com/help/matlab/%
ref/ind2sub.html]{ind2sub()} functions.

As the names may suggest, they return return one indexing regime given
the other.  The only requirement for these functions is the size of the
array you're working with:

\mCommandOutput{03-indexing.d/ind2sub}

\mCommandOutput{03-indexing.d/sub2ind}

\section{Logical Arrays}

We've finally reached the point where we can introduce a new datatype:
the \emph{logical} datatype.  MATLAB logical values operate over
\(\Z_2\) where \mCommand{0} denotes \mCommand[https://www.mathworks.%
com/help/matlab/ref/false.html]{false}, and \mCommand{1} denotes
\mCommand[https://www.mathworks.com/help/matlab/ref/true.html]{true}.

What makes logical arrays so powerful is that they obey the rules of
Boolean Algebra and allow for us to extract elements of arrays.

\begin{longtable}{p{0.15\textwidth}p{0.65\textwidth}}
\toprule
Operator & Purpose \\
\midrule
\mCommand{&}   & Logical AND                                         \\
\mCommand{&&}  & Logical AND with short-circuiting                   \\
\mCommand{~}   & Logical NOT                                         \\
\mCommand{|}   & Logical OR                                          \\
\mCommand{||}  & Logical OR with short-circuiting                    \\
\mCommand{xor} & Logical exclusive-OR                                \\
\mCommand{all} & Determine if all array elements are nonzero or true \\
\mCommand{any} & Determine if any array elements are nonzero         \\
\bottomrule
\\
\caption{Logical Operations}
\end{longtable}

Say we wanted to extract the corners of \(\mat{A}\).  We could create a
logical array with \mCommand[https://www.mathworks.com/help/matlab/ref/%
logical.html]{logical()} with \(1\)'s in the location of the elements
we'd like to extract:

\inputminted{matlab}{03-indexing.d/corners.m}

\begin{minipage}{\textwidth}
	We can then use this logical array to index into \(\mat{A}\):

	\vspace{1em}

	\mCommandOutput{03-indexing.d/select}
\end{minipage}

\section{Relational Operations}

Now that we know about logical values, we can combine this power with
relational operators to select elements satisfying a condition.

\begin{longtable}{p{0.15\textwidth}p{0.55\textwidth}}
\toprule
Operator & Purpose \\
\midrule
\mCommand{==}       & Equality                                                \\
\mCommand{>=}       & Greater than or equal to                                \\
\mCommand{>}        & Greater than                                            \\
\mCommand{<=}       & Less than or equal to                                   \\
\mCommand{<}        & Less than                                               \\
\mCommand{~=}       & Inequality                                              \\
\mCommand{isequal}  & Array equality                                          \\
\mCommand{isequaln} & Array equality, treating \mCommand{NaN} values as equal \\
\bottomrule
\\
\caption{Relational Operations}
\end{longtable}

For example:

\mCommandOutput{03-indexing.d/lt}

Notice how the result of this operation is a logical array, which might
tempt you to utilize our previously learned logical operators!

Say we wanted to get all values in the domain \((3, 7)\):

\mCommandOutput{03-indexing.d/domain}

As another example, we can use the \mCommand[https://www.mathworks.com/%
help/matlab/ref/find.html]{find()} function to get the linear index of
the element(s) that equal 5:

\mCommandOutput{03-indexing.d/find}

As a final example, let's select all the columns of \(\mat{A}\) that
sum to a value greater than 30:

\mCommandOutput{03-indexing.d/satisfy}

Oops, it looks like we didn't get what we wanted!  Unfortunately, MATLAB
doesn't broadcast when indexing, so we'll need to use \mCommand[https:%
//www.mathworks.com/help/matlab/ref/repmat.html]{repmat()} to generate
the logical matrix we desire and a \mCommand[https://www.mathworks.com/%
help/matlab/ref/reshape.html]{reshape()} to get the final output in the
correct format.

\mCommandOutput{03-indexing.d/cols}

As you can see, these operations are \emph{powerful} but may require you
to massage the outputs to get it into the correct format.  At first,
they may seem unintuitive, but like all things in life, you'll get the
hang of it with persistence and patience.
\end{document}
