% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 05-plotting.tex -- my pretty data visualizations
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Lesson 05: Plotting}
\input{postamble}

\begin{document}
\renderTitle

We've toyed a lot with how to perform calculations, but it's not awfully
useful if we can't visualize them.  Thankfully, MATLAB offers us a wide
array of plotting functionality to help us visualize our data.

\section{Graphic Objects}

MATLAB organizes graphical objects into a hierarchical structure%
\footurl{https://www.mathworks.com/help/matlab/graphics-objects.html}:

\begin{figure}[!ht]
\Centering
\small
\begin{tabular}{c}
	\lstinputlisting[basicstyle=\ttfamily]%
	{05-plotting.d/graphics-object-hierarchy.txt}
\end{tabular}
\caption{Organization of Graphics Objects}
\end{figure}

\section{2-D Plots}

There's not much point in explaining.  Let's jump right into an example.

\mCommandGraphic{05-plotting.d/basic-plot}

Here \mCommand[https://www.mathworks.com/help/matlab/ref/figure.html%
]{figure} creates a new figure that \mCommand[https://www.mathworks.%
com/help/matlab/ref/plot.html]{plot} can interact with.  Although we
successfully plotted our sine wave, it's quite ugly.  Let's take the
liberty to make it a little more pretty.

\mCommandGraphic{05-plotting.d/pretty-sin}

\subsection{Retaining Plots}

Say we wanted to plot multiple functions on the same plot:

\aside{\mCommand{hold on} can be a back stabber!  Sometimes you may
experience buggy behavior if you call it before the first plot.}

\mCommandGraphic{05-plotting.d/retain}

\subsection{Subplots}

We can have multiple plots in a figure by doing the following:

\aside{Notice the zero-crossing property of the \(\symrm{sinc}\)
function!}

\mCommandGraphic{05-plotting.d/subplots}

\section{3-D Plots}

MATLAB is also capable of making 3-D plots:

\mCommandGraphic{05-plotting.d/sinc-surf}

You can think of \mCommand[https://www.mathworks.com/help/matlab/ref/%
meshgrid.html]{meshgrid} as the Cartesian product of the two input
vectors split over two output matrices.  Having these matrices allows us
to evaluate functions over a 2-D plane.

\newpage

We can also use a mesh instead:

\mCommandGraphic{05-plotting.d/sinc-mesh}

\section{More Plotting Functionality}

This barely scratches the surface of MATLAB's extensive plotting
functionality.  Once again, the online documentation is a great starting
point for the interested reader: \url{https://www.mathworks.com/help/%
matlab/2-and-3d-plots.html}
\end{document}
