% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 09-fixed-point-arithmetic.tex -- precision? never head of her
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Lesson 09: Fixed-Point Arithmetic}
\input{postamble}

\begin{document}
\renderTitle

One day, you may have the great fortune of implementing a digital filter
on an embedded device where you only have a few microseconds to filter
before the next sample.  In such a scenario, performing floating-point
operations is impractical due to how much slower they are to \vocab{%
Multiply–Accumulate (MAC)} operations.  To benefit from MAC operations,
we turn to fixed-point arithmetic, but in the process, we open Pandora's
box and must account for the substantial quantization erorr we've
introduced.  Luckily for us, MATLAB offers us a Fixed-Point Designer%
\footurl{https://www.mathworks.com/help/fixedpoint/index.html} for
optimizing and implementing fixed-point algorithms.

\section{Constructing Fixed-Point Types}

\begin{figure}[!ht]
\Centering
\small
\begin{tabular}{c}
	\lstinputlisting[basicstyle=\ttfamily]%
	{09-fixed-point-arithmetic.d/fixed-point.txt}
\end{tabular}
\caption{Fixed-Point Value}
\end{figure}

First, we must tell MATLAB the type of fixed-point value we wish to
construct:

\mCommandOutput{09-fixed-point-arithmetic.d/fixed-point-type}

In our case, we've constructed a signed 32-bit fixed-point value with a
fraction length of 30 bits, giving us a range from -2 to \(2 - 2^{-30}\)
in increments of \(2^{-30}\).

Now, to convert a \mCommand{double} to a fixed-point value, we call
\mCommand[https://www.mathworks.com/help/fixedpoint/ref/embedded.fi.%
html]{fi()}:

\mCommandOutput{09-fixed-point-arithmetic.d/fi-call}

\subsection{Set Fixed-Point Math Settings}

At times, we may wish to change MATLAB's default behavior when
performing fixed-point math.  Say we wanted to have two's complement
overflow as opposed to saturation:

\mCommandOutput{09-fixed-point-arithmetic.d/fi-saturate}

\mCommandOutput{09-fixed-point-arithmetic.d/fi-wrap}

We can also apply these properties to new fixed-point types with
\mCommand[https://www.mathworks.com/help/fixedpoint/ref/embedded.%
fimath.html]{fimath()}:

\mCommandOutput{09-fixed-point-arithmetic.d/fimath-OverflowMode}

\section{Fixed-Point Filter Design}

For this part of the lesson, we'll be working with the brake pressure
sensor data from Assignment
07:\footurl{https://raw.githubusercontent.com/jacobkoziej/jk-ece210/master/src/assignments/07-under-pressure.d/40p_1000ms.csv}

\inputminted{matlab}{09-fixed-point-arithmetic.d/load-data.m}

Since fixed-point values are objects, we can't use our standard
filtering functions on them, so we turn to filter objects:

\aside{We use the Direct Form I structure over Direct Form II to gain
more numerical stability at the cost of using double the amount of state
variables.}

\inputminted{matlab}{09-fixed-point-arithmetic.d/lpf.m}

We can then apply the filter to our quantized data like so:

\mCommandGraphicSplit{09-fixed-point-arithmetic.d/lpf-apply}

Let's half our bit allocation to see how our filter performs:

\mCommandGraphicSplit{09-fixed-point-arithmetic.d/lpf-poor}

Ouch, it looks like the quantization error was too much to handle!  To
combat this, we can decrease our filter's requirements to avoid higher
orders and hence numerical instability in the form of overflows, but
keep in mind this isn't a full-proof solution.
\end{document}
