% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% 04-func-off.tex -- i'm not doing this again
% Copyright (C) 2024  Jacob Koziej <jacobkoziej@gmail.com>

\documentclass{article}

\input{preamble}
\title{Assignment 04: Func Off}
\input{postamble}

\begin{document}
\renderTitle

\begin{enumerate}[leftmargin=*]
	\item
		Create an \mCommand{ip} anonymous function that performs
		the standard inner product over \(\C\) and an
		\mCommand{ip_norm} anonymous function for its associated
		\(\symrm{L}^2\) norm.

	\item
		Create a \mCommand{gram_schmidt} function.  The input
		should be a matrix of linearly independent columns and
		function handles that define an inner product and norm
		for the Gram–Schmidt process.  The function should
		return a matrix of orthonormal column vectors.

	\item
		Create an \mCommand{isorthogonal} function that accepts
		two vectors and an inner product function handle.  The
		function should return \mCommand{true} if the vectors
		are orthogonal and \mCommand{false} if they are not.
		Due to the numerical instability of the Gram-Schmidt
		process, our orthonormal vectors may not be exactly
		orthogonal.  Take this into account by utilizing the
		\mCommand[https://www.mathworks.com/help/matlab/ref/%
		eps.html]{eps} function.

	\item
		Calculate orthonormal vectors from the following
		orthogonal set:
		\begin{equation}
			\symcal{S}
			=
			\left\{
				\begin{bmatrix}
					1       \\
					\im     \\
					2 - \im \\
					-1      \\
				\end{bmatrix},
				\begin{bmatrix}
					2 + 3\im \\
					3\im     \\
					1 - \im  \\
					2\im     \\
				\end{bmatrix},
				\begin{bmatrix}
					-1 + 7\im \\
					6 + 10\im \\
					11 - 4\im \\
					3 + 4\im  \\
				\end{bmatrix}
			\right\}
		\end{equation}
		and store them into matrix \mCommand{U}.
\end{enumerate}
\end{document}
